---
title: "Resampling KTR ETAs"
author: "Jim Hughes"
date: "2 November 2017"
output:
  html_document: default
  html_notebook: default
---
```{r setup, include = FALSE}
# Load libraries
  library(knitr)

# Load in data
  fit_dir <- "E:/Hughes/Data/PK/FLAG/COV15/RUN016_CL_CRCL2_FFM.nm7"
  fitdata <- read.csv(paste0(fit_dir, "/run016_cl_crcl2_ffm.fit.csv"))
  
  sim_dir <- "E:/Hughes/Data/PK/FLAG/COV15/RUN016_CL_CRCL2_FFM_VPC.nm7"
  simdata <- read.csv(paste0(sim_dir, "/RUN016_CL_CRCL2_FFM_VPC.fit.csv"))
  
# Create summary statistic function
  sumfuncLOTS <- function(x) {
    stat1 <-  mean(x, na.rm=T)
    stat2 <-  sd(x, na.rm=T)
    stat3 <-  min(x, na.rm=T)
    stat4 <-  max(x, na.rm=T)
    stat5 <-  length(x)
    stat6 <-  quantile(x, probs=0.05, na.rm=T, names=F)  #90%CI
    stat7 <-  median(x, na.rm=T)
    stat8 <-  quantile(x, probs=0.95, na.rm=T, names=F)
    result <- c("mean"=stat1,"sd"=stat2, "min"=stat3, 
      "CI90lo"=stat6, "median"=stat7, "CI90hi"=stat8, "max"=stat4, "n"=stat5)
    result
  }
```
Due to food effect the Tmax of lenalidomide can differ greatly depending on the 
patient. No control was done for the food effect in clinical trials, therefore 
the between subject variability on absorption must compensate. The large %CV on 
KTR along with the transit model is resulting in good fits, but poor 
simulations. The assumption is that the poor simulations are due to ETAs on the 
low end creating patients that cannot exist.

### Literature
Chen et. al's _Single-dose pharmacokinetics of lenalidomide in healthy 
volunteers: dose proportionality, food effect, and racial sensitivity_ looked 
at the effect of food on lenalidomide. They found that when compared to fasting 
patients, a high fat, high calorie meal caused:

* Cmax to reduce by 50% when taken with high fat meal
* AUC to reduce by 20% when taken with high fat meal
* Tmax to increase by 1.63 hours when taken with high fat meal

This was determined using 17 healthy individuals in a 1 phase crossover trial. 
The median fasting tmax was 0.75 hours and the median fed tmax was 2.5 hours. 
The median difference between the the fasting and fed state for the 17 patients 
was 1.63 hours. The 90% confidence intervals were 1.38 - 2 hours.

### Study Data
So lets see if the study data agrees with these values.

Initially determine the minimum and maximum KTRs fitted by the final model when 
used on the study data.

```{r fit_minmaxKTR, eval = T}
  round(sumfuncLOTS(fitdata$KTR), digits = 2)
```

Next lets look at these extreme individuals

```{r fit_extremeID, eval = T}
  mindata <- fitdata[which(fitdata$KTR == min(fitdata$KTR)),]
  kable(mindata[c(2:7, 26:28, 31)], digits = 2)
  maxdata <- fitdata[which(fitdata$KTR == max(fitdata$KTR)),]
  kable(maxdata[c(2:7, 26:28, 31)], digits = 2)
  maxtmax <- mindata$TIME[which(mindata$DV == max(mindata$DV))]
  mintmax <- maxdata$TIME[which(maxdata$DV == max(maxdata$DV))]
  c(min = mintmax, max = maxtmax)
```

The Tmax's here range from 0.75 to 4 hours. This is a bit wider than seen in 
Chen et. al's paper. But these are patients, not healthy young males, so it is 
possible than in some of these patients absorption times could be longer due to 
aging, disease and other factors.

### Simulation Data
So how bad is the simulation data? Or is our initial assumption wrong?

Initially determine summary statistics from simulation KTR values.

```{r sim_minmaxKTR, eval = T}
  round(sumfuncLOTS(simdata$KTR), digits = 2)
```

The 90% confidence intervals, mean and median seem alright, but the minimum and 
maximum are really off. Probably good to find out the tmax associated with those
KTRs.

```{r sim_extremeID, eval = T}
  mindata <- simdata[which(simdata$KTR == min(simdata$KTR)),]
  kable(mindata[c(2:7, 26:28, 31)], digits = 2)
  maxdata <- simdata[which(simdata$KTR == max(simdata$KTR)),]
  kable(maxdata[c(2:7, 26:28, 31)], digits = 2)
  maxtmax <- mindata$TIME[which(mindata$DV == max(mindata$DV))]
  mintmax <- maxdata$TIME[which(maxdata$DV == max(maxdata$DV))]
  c(min = mintmax, max = maxtmax)
```

So the minimum is 0.25 hours, which does not mimic reality at all. The maximum 
is 32 hours, which is ridiculous!

Despite this the 90% confidence intervals are fine, which is good! But it might 
mean that resampling the ETAs will not improve VPCs...

Should resample ETA's to ensure that KTR is within the boundaries 
seen in the Study data. Therefore resample if KTR less than 3 or KTR more than 40

### Outcome
Did not work out, VPCs do not improve substantially, but a nice exploration of 
how to truncate distributions in NONMEM!
