---
title: "Protein Binding Calculations for Equilibrium Dialysis"
author: "Jim Hughes"
date: "Tuesday, 10th October 2017"
output:
  html_document: default
---
```{r setup, include = FALSE}
# Set up directories
  if (!exists("git.dir")) {
    rm(list = ls(all = T))
    wd <- c("C:/Users/Jim Hughes/Documents", "C:/Users/hugjh001/Documents",
      "C:/Users/hugjh001/Desktop", "C:/windows/system32")

    graphics.off()
    if (getwd() == wd[1]) {
      gir.dir <- paste0(getwd(), "/GitRepos")
      reponame <- "len_pbpk"
    } else if (getwd() == wd[2]) {
      git.dir <- getwd()
      reponame <- "len_pbpk"
    } else if (getwd() == wd[3] | getwd() == wd[4]) {
      git.dir <- "E:/Hughes/Git"
      reponame <- "len_pbpk"
    }
    rm("wd")
  }

# Load libraries
  library(readxl)
  library(stringr)
  library(plyr)
  library(knitr)

# Load in data
  data_folder <- "E:/Hughes/Data/RAW_NonClinical/protein_binding/20171009"
  filename_in <- paste0(data_folder, "/mouse plasma protein binding results_9282017_std passing_Short.xls")
  mouse_plas <- read_excel(filename_in, sheet = "Lenalidomide", skip = 4)

  filename_in <- paste0(data_folder, "/Plasma protein binding_plasma_human_rerun_finalresults_9282017_Short.xls")
  human_plas <- read_excel(filename_in, sheet = "Lenalidomide", skip = 4)

  filename_in <- paste0(data_folder, "/Plasmaproteinbinding_PBS_results_10022017_Short_171006133211.xls")
  both_pbs <- read_excel(filename_in, sheet = "Lenalidomide", skip = 4)

# Combine data
  any(names(mouse_plas) != names(human_plas) & names(mouse_plas) != names(both_pbs))
  alldata <- rbind(mouse_plas, human_plas, both_pbs)

# Fix column names
  name_correction <- c("Specified.Amount", "Calculated.Amount")
  names(alldata) <- str_replace_all(names(alldata),"[ ()#]",".")
  names(alldata)[names(alldata) == "Amount"] <- name_correction

# Take subset of sample data and add covariate info
  subdata <- alldata[alldata$Sample.Type == "Unknown Sample" & !is.na(alldata$Sample.Type),]

  subdata$Species <- "mouse"
  subdata$Species[str_detect(subdata$Filename, "h")] <- "human"

  subdata$Vehicle <- "plasma"
  subdata$Vehicle[str_detect(subdata$Filename, "b")] <- "buffer"

  subdata$id <- 4
  subdata$id[str_detect(subdata$Filename, "5")] <- 5
  subdata$id[str_detect(subdata$Filename, "6")] <- 6

  subdata$Conc <- 0.3
  subdata$Conc[str_detect(subdata$Filename, "0_03")] <- 0.03
  subdata$Conc[str_detect(subdata$Filename, "10_")] <- 10

  subdata <- with(subdata, data.frame(
    ID = id,
    name = Filename,
    dv = suppressWarnings(as.numeric(Calculated.Amount)),
    species = Species,
    vehicle = Vehicle,
    conc = Conc,
    peak_status = Peak.Status,
    area = as.numeric(Area),
    area_istd = as.numeric(ISTD.Area),
    area_ratio = as.numeric(Area.Ratio),
    rt = as.numeric(RT)
  ))
  
# Handle negatives then split into human and mouse datasets
  subdata$dv[subdata$dv <= 0.3] <- NA
```

The Boudinot formula was used to determine protein binding:
$$fb = \frac{(C_{p}-C_{d})V_{pe}/V_{pi}}{\left[(C_{p}-C_{d})V_{pe}/V_{pi}\right]+C_{d}}$$
where 

* $C_{p}$ is the concentration on the sample side (plasma)
* $C_{d}$ is the concentration on the dialysate side (buffer)
* $V_{pe}$ is the equilibrium plasma volume
* $V_{pi}$ is the initial plasma volume

When $V_{pe}$ is equal to $V_{pi}$ the equation is simplified to
$$fb = \frac{C_{p}-C_{d}}{C_{p}}$$

This was applied using the function below, using it on each unique set of 
species, concentration and ID
```{r function}
  fb_fun <- function(x) {
    Cp <- x$dv[x$vehicle == "plasma"]  # total plasma conc
    Cd <- x$dv[x$vehicle == "buffer"]  # free dialysate conc
    Ci <- unique(x$conc)*1000  # initial plasma conc
    fb <- (Cp-Cd)/Cp*100  # fraction bound
    dr <- Cd/Ci*100  # drug recovered in dialysate
    pr <- Cp/Ci*100  # drug recovered in plasma
    c(fraction_bound = fb, plas_recovered = pr, dial_recovered = dr, total_recovered = dr+pr)
  }
```

As an example a specific subset of data is run through the processes involved
in the function below

```{r example}
  example <- subset(subdata, species == "human" & conc == 10 & ID == 4)
  kable(example, digits = 2, caption = "Subsetted data")
  plasma_conc <- example$dv[example$vehicle == "plasma"]
  buffer_conc <- example$dv[example$vehicle == "buffer"]
  delta_conc <- plasma_conc - buffer_conc
  c("plasma" = plasma_conc, "buffer" = buffer_conc, "dC" = delta_conc)
  delta_conc/plasma_conc*100
  fb_fun(example)["fraction_bound"]
```

When applied to each set of samples the result is as seen below

```{r output}
  kable(ddply(subdata, .(species, conc, ID), fb_fun), digits = 2)
```

### Data Management

Before analysis all data is combined into one data.frame. The methods are
described below:

* The three spreadsheets are read in using `readxl::read_excel` skipping the
first 4 lines
* The column names are standardised, with the two "Amount" columns being
corrected to Specifiied and Calculated amounts
* The sample data is subset from the rest of the data using the Sample Type
column as they are all marked with "Unknown Sample"
* Sample information about species, concentration, vehicle (plasma vs. buffer)
and id number are extracted from Filename column
* All values less than or equal to the LLOQ were set to NA

The resulting data.frame is shown below:
```{r data}
  kable(subdata, digits = 2)
```
