---
title: "Data check and tidy for IV mouse data"
author: "Jim Hughes"
date: "Wednesday, 5 July 2017"
output:
  html_notebook: default
  html_document: default
---
Data originally published:
_Rozewski DM, Herman SEM, Towns WH, Mahoney E, Stefanovski MR, Shin JD, et al._
_Pharmacokinetics and Tissue Disposition of Lenalidomide in Mice. AAPS J._
_2012;14(4):872-82._

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)

if (!exists("git.dir")) {
  rm(list=ls(all = TRUE))
  graphics.off()
  #git.dir <- "E:/Hughes/Git"
  git.dir <- "C:/Users/Jim Hughes/Documents/GitRepos"
  reponame <- "len_pbpk"
}

source(paste(git.dir, reponame, "rscripts",
  "data_load.R", sep = "/"))
s
scriptname <- "datacheck_iv"
```
# --------------------------------------------------------------------------------
### Explore Data for Cleaning and Data Extraction
#### View column names and structure
```{r}
names(rawiv)
str(rawiv)
```


#### Check distribution of doses and weights
```{r}
  with(rawiv, table(Dose..mg.kg., useNA = "always"))
  with(rawiv, table(Mouse.Wt..g., useNA = "always"))
```


#### Check NA's for non-binned data
```{r}
any(is.na(rawiv$Sample.ID))
any(is.na(rawiv$Time..min.))
```


There are NA's in the time column, this could be problematic.
Identify where the NA's are located
```{r, results='asis'}
rawiv[which(is.na(rawiv$Dose..mg.kg.)), ]
rawiv[which(is.na(rawiv$Mouse.Wt..g.)), ]
rawiv[which(is.na(rawiv$Time..min.)), ]
```


NA's appear to be as a result of a repeat sample, it was ascertained from 
correspondence with Mitch that these are in fact for the same mouse.
These data entries will need to be combined into a single data entry.

#### Count number of samples (including repeats mentioned above)
```{r}
length(unique(rawiv$Sample.ID))
```


### Clean & Extract Data
Data converted to following units:

* Dose - mg
* AMT - ng
* Weight - g
* Concentrations - ng/mL
* Time - min
```{r, cache = T, echo = F}
dataiv <- data.frame("ID" = rawiv$Sample.ID, "DOSEMGKG" = rawiv$Dose..mg.kg.)
dataiv$DOSEMG <- dataiv$DOSEMGKG*rawiv$Mouse.Wt..g./1000
dataiv$AMT <- dataiv$DOSEMG*10^6  #dose in ng
dataiv$WT <- rawiv$Mouse.Wt..g.
dataiv$TIME <- rawiv$Time..min.
dataiv$PLA <- rawiv$Plasma.DV..ng.mL.
dataiv$BRA <- rawiv$Brain..ng.mL.
dataiv$LVR <- rawiv$Liver..ng.mL.
dataiv$MSC <- rawiv$Mscl..ng.mL.
dataiv$HRT <- rawiv$Hrt..ng.mL.
dataiv$SPL <- rawiv$Spln..ng.mL.
dataiv$LUN <- rawiv$Lung..ng.mL.
dataiv$KID <- rawiv$Kidney..ng.mL.
```


Sample IDs cleaned up using function that splits the ID string into ID and 
nominal time after dose (TADNOM).
```{r, cache = T}
head(dataiv)
IDiv <- end.splitter(dataiv$ID)
dataiv <- cbind(IDiv, dataiv[,-1])
head(dataiv)
```


#### Check the nominal time against dose.
```{r}
with(dataiv, table(DOSEMGKG, TADNOM))
dataiv[which(dataiv$TADNOM == "25"), ]
```


There are one set of values stated to have TADNOM as 25, but values were
measured at 20mins, one at 13 which is odd.
The TADNOM is likely meant to be 20, and since this data will only be used for
categorising the data the TADNOM was changed to 20 for these rows.
```{r, cache = T, echo = F}
dataiv$TADNOM[which(dataiv$TADNOM == "25")] <- "20"
```


The string splitting function identify the repeated sample with a repeated UID.
There are two lots of concentrations in these rows, so which one is correct?
```{r, cache = T}
subiv <- dataiv[which(with(dataiv, TADNOM == "20" & DOSEMGKG == 5 & !is.na(dataiv$WT))), ]
IDori <- dataiv[which(is.na(dataiv$WT)), ]
IDrep <- arrange(dataiv[which(is.na(dataiv$DOSEMGKG)), ], UID)
tissue.mean <- function(x) {
  colwise(mean)(x[c("PLA", "BRA", "LVR", "MSC", "HRT", "SPL", "LUN", "KID")])
}
rbind(tissue.mean(subiv), tissue.mean(IDori), tissue.mean(IDrep))
```


The first row represents the mean of all data within the same time and dose 
bracket, excluding the repeated samples. The second row represents the samples
with DOSEMGKG values and the third row represents the samples with WT & TIME 
values.

It appears that the row with WT & TIME has no usable concentrations. These rows
were combined using the dependent variables and dose from one and the weight
and time from the second.

The data should now be cleaned up and usable

```{r, cache = T, echo = F}
# Combine into one row, use DV and DOSEMGKG from orig, use WT & TIME from repeat
IDori$TIME <- IDrep$TIME
IDori$WT <- IDrep$WT
IDori$DOSEMG <- IDori$DOSEMGKG*IDori$WT/1000
IDori$AMT <- IDori$DOSEMG*10^6  # units: ng
IDori$PLA <- NA

# Then replace both samples with this combined sample and order the data.frame
IDrem1 <- which(is.na(dataiv$WT))
IDrem2 <- which(is.na(dataiv$DOSEMGKG))
dataiv <- arrange(
  rbind(
    dataiv[-c(IDrem1, IDrem2), ], IDori
  ),  # rbind
  DOSEMGKG, TIME
)  # arrange

# Make TADNOM numeric
dataiv$TADNOM <- as.numeric(dataiv$TADNOM)

# Create data.frame with average values for each timeslot
dataiv.av <- ddply(dataiv, .(DOSEMGKG, TADNOM), function(x) {
  data.frame(
    "DOSEMG" = mean(x$DOSEMG, na.rm = T),
    "AMT" = mean(x$AMT, na.rm = T),
    "WT" = mean(x$WT, na.rm = T),
    "TIME" = mean(x$TIME, na.rm = T),
    "PLA" = mean(x$PLA, na.rm = T),
    "BRA" = mean(x$BRA, na.rm = T),
    "LVR" = mean(x$LVR, na.rm = T),
    "MSC" = mean(x$MSC, na.rm = T),
    "HRT" = mean(x$HRT, na.rm = T),
    "SPL" = mean(x$SPL, na.rm = T),
    "LUN" = mean(x$LUN, na.rm = T),
    "KID" = mean(x$KID, na.rm = T)
  )
})
```


### Further Data Checking
#### Check subject numbers
Also whether repeat sample was successfully removed.
```{r}
with(dataiv, table(ID))
with(dataiv, table(UID))
any(with(dataiv, table(UID)) > 2)
```


#### Check PK dose data
```{r}
with(dataiv, table(DOSEMGKG, TADNOM))
```


#### Check the dose columns
```{r}
with(dataiv, table(DOSEMG*1000))
hist(dataiv$DOSEMG*1000)
```


#### Check the distribution of DV
For all data and for the mean data
```{r, cache = T}
meltiv <- melt(dataiv, id = c("UID", "ID", "TADNOM", "DOSEMGKG", "DOSEMG", "AMT", "WT", "TIME"))
colnames(meltiv) <- c(head(colnames(meltiv), 8), "TISSUE", "DV")
iv.distplot(meltiv, "alldata", NULL)
meltiv.av <- melt(dataiv.av, id = c("DOSEMGKG","TADNOM", "DOSEMG", "AMT", "WT", "TIME"))
colnames(meltiv.av) <- c(head(colnames(meltiv.av), 6), "TISSUE", "DV")
iv.distplot(meltiv.av, "meandata", NULL)
```

```{r, cache = T, echo = F}
meltiv$DVNORM <- meltiv$DV/meltiv$DOSEMG
meltiv$MDV <- ifelse(is.na(meltiv$DV), 1, 0)
meltiv.av$DVNORM <- meltiv.av$DV/meltiv.av$DOSEMG
meltiv.av$MDV <- ifelse(is.na(meltiv.av$DV), 1, 0)
```


#### Plot PK data
```{r}
plot.out <- NULL
iv.CvTplot(meltiv[meltiv$DOSEMGKG == 0.5,], "alldata", plot.out)
iv.CvTplot(meltiv[meltiv$DOSEMGKG == 1.5,], "alldata", plot.out)
iv.CvTplot(meltiv[meltiv$DOSEMGKG == 5,], "alldata", plot.out)
iv.CvTplot(meltiv[meltiv$DOSEMGKG == 10,], "alldata", plot.out)
iv.CvTplot(meltiv.av[meltiv.av$DOSEMGKG == 0.5,], "meandata", plot.out)
iv.CvTplot(meltiv.av[meltiv.av$DOSEMGKG == 1.5,], "meandata", plot.out)
iv.CvTplot(meltiv.av[meltiv.av$DOSEMGKG == 5,], "meandata", plot.out)
iv.CvTplot(meltiv.av[meltiv.av$DOSEMGKG == 10,], "meandata", plot.out)
iv.CvTplot(meltiv, "alldata", plot.out, dosenorm = T)
iv.CvTplot(meltiv.av, "meandata", plot.out, dosenorm = T)
```


### Covariate Data Check
#### Count missing by dose group
```{r}
ddply(dataiv, .(DOSEMGKG), colwise(calculate.percent.missing))
```


#### Count missing by nominal time after dose
```{r}
ddply(dataiv, .(TADNOM), colwise(calculate.percent.missing))
```


#### Summary of missing data
```{r}
ddply(dataiv, .(DOSEMGKG, TADNOM), colwise(calculate.percent.missing))
```


#### DV count
```{r, cache = T, echo = F}
DVsum <- function(data, group, uid, col) {
    ind <- unique(data[, group])
    out <- data.frame(
      matrix(nrow = length(ind), ncol = length(col)*2 + 1)
    )
    out[1] <- ind
    for (i in 1:length(ind)) {
      sub <- data[data[group] == ind[i], ]
      out[i, 2] <- length(unique(sub[, uid]))
      for (j in 1:length(col)) {
        out[i, j*2 + 1] <- length(which(!is.na(sub[, col[j]])))
        out[i, j*2 + 2] <- out[i, j*2 + 1]/out[i, 2]
      }
    }
    names(out) <- c(
      group, "SUBcount", paste0(rep(col, each = 2), c("count", "perSUB"))
    )
    return(out)
  }
```

```{r}
DVsum(dataiv, group = "DOSEMGKG", uid =  "UID",
  col = c("AMT", "PLA", "BRA", "LVR", "MSC", "HRT", "SPL", "LUN", "KID")
)
```

```{r, cache = T, echo = F}
plotdata <- meltiv
BINnumber <- 3

plotdata$DOSEMGKGf <- as.factor(plotdata$DOSEMGKG)
plotdata$TISSUEf <- as.factor(plotdata$TISSUE)
plotdata$WT_bin <- as.factor(ave(plotdata$WT, cut(plotdata$WT, BINnumber), FUN = median))
plotdata$DOSE_bin <- as.factor(ave(plotdata$DOSEMG, cut(plotdata$DOSEMG, BINnumber), FUN = median))
```


#### Plot by Factor
```{r, cache = T}
plotByFactor("DOSE_bin", "Binned Dose (ug)", plotdata, plot.out)
plotByFactor("WT_bin", "Binned Weight (g)", plotdata, plot.out)
```


