---
title: "Learning about Fourier Curves"
output: 
  html_document: default
  word_document: default
---
_A description of how fourier curves work and an exploration into how these_
_can be used to fit curves to pharmacokinetic absorption data._
_For more information check out these links:_

https://onlinelibrary.wiley.com/doi/pdf/10.1002/9781118593233.ch12

http://www.di.fc.ul.pt/~jpn/r/fourier/fourier.html

# --------------------------------------------------------------------------------

```{r sim.data, echo = F}
# Setup workspace
  suppressMessages(library(ggplot2))
  
  theme_bw2 <- theme_set(theme_bw(base_size = 14))
  theme_update(plot.title = element_text(hjust = 0.5))

# Absorption Curve
  time.samp <- seq(0, 24, by = 1)
  absdata <- data.frame(
    time = time.samp,
    line1 = -0.4*time.samp + 4,
    line2 = -0.2*time.samp + 4
  )
  absdata$sumexp <- exp(absdata$line2) - exp(absdata$line1)
  subdata <- data.frame(time = c(0, 1, 2, 4, 6, 8, 12, 16, 24))
  subdata$dv <- absdata$sumexp[which(absdata$time %in% subdata$time)]
```

### Simple Sinusoidal Waves

```{r wave1}
  xs <- seq(-2*pi, 2*pi, pi/100)
  wave.1 <- sin(3*xs)
  wave.2 <- sin(10*xs)
  plot(xs,wave.1, type = "l", ylim = c(-1, 1))
  abline(h = 0,lty = 3)
  plot(xs,wave.2, type = "l", ylim = c(-1, 1))
  abline(h = 0, lty = 3)
```

Here we have two sine waves. Sine wave notation is below:
$$y(t) = Asin(2\pi ft+\phi) = Asin(\omega t+\phi)$$
where:

* $A$ is the amplitude (peak deviation from zero)

* $f$ is the frequency, the number of oscillations that occur each unit of time

* $\omega$ is angular frequency, rate of change of the function in units of 
radians per unit of time

* $\phi$ is the phase, specifies in radians where in the cycle the oscillation 
is at $t = 0$

It should be noted that a phase shifted sine wave can also be represented as 
seen below:
$$sin(a+b)=sin(a)cos(b)+cos(a)sin(b)$$

### Fourier Series using Sine Waves
Joseph Fourier showed that any periodic wave can be represented by a sum of 
simple sine waves. This sum is called the **Fourier Series**.

```{r wave2}
  wave.3 <- 0.5 * wave.1 + 0.25 * wave.2
  plot(xs, wave.3, type = "l")
  title("Fourier Series")
  abline(h = 0,lty = 3)
  repeat.xs <- seq(-2*pi, 0, pi/100)
  wave.3.repeat <- 0.5*sin(3*repeat.xs) + 0.25*sin(10*repeat.xs)
  points(repeat.xs, wave.3.repeat, type = "l",col = "red")
  abline(h = 0,v = c(-2*pi, 0), lty = 3)
```

A Fourier Series is periodic, repeating according to it's frequency $f$ and 
wavelength $\lambda$. Non-periodic waves do not have a frequency or wavelength.
What occurs within a single period is more interesting for pharmacokinetics.

Some important things are:

* The **fundamental period**, $T$, is the period of time that the samples were 
taken in. For the case of an oral drug this could be 24 hours. If it is 
multi-dose data then the fundamental period is 24 multiplied by the number of 
sequential doses.

* The **sampling rate**, $sr$ is the number of samples taken over a time period. 
The interval between these times is called the **sample interval**, $si$, 
which is the fundamental period time divided by the number of samples $N$. 
Therefore $si = \frac{T}{N}$.

* The **fundamental frequency**, $f_0$, which is $\frac{1}{T}$. The fundamental
frequency of the repeating pattern, or how long the wavelength is. In the waves
above, the fundamental frequency was $\frac{1}{2\pi}$. The frequencies of the
wave components must be integer multiples of the fundamental frequency.

* The **angular frequency**, mentioned in the description of a sine wave is 
determined from these values like so, $\omega=2\pi f_0$

The Fourier Series above is represented as:
$$f(t)=0.5sin(3\omega t)+0.25sin(10\omega t)$$

It should be noted that the frequency $f$ of the simple waves, is separate from 
the fundamental frequency $f_0$ of the complex wave, which is used to determine 
the angular frequency $\omega$.

So if we were to implement this as a function in R:

``` {r fourierfn}
# Set up data parameters
  acq.f <- 100  # data aquisition frequency
  t.int <- 6  # measuring time interval
  t <- seq(0, t.int, 1/acq.f)  # vector of sampling times
  f.0 <- 1/t.int  # fundamental frequency
  w <- 2*pi*f.0  # angular frequency
  
# Set up wave parameters
  in.freq <- c(3, 10)  # frequency of sine waves
  in.phase <- c(0, 0)  # phase (delay of sine waves)
  in.amp <- c(0.5, 0.25)  # amplitude (strength of sine waves)
  
# Define function
  fn <- function(t, w) {
    sum(in.amp*sin(in.freq*w*t + in.phase))
  }
  
# Plot using this function
  pred <- sapply(t, function(t) fn(t, w))
  plot(t, pred, type = "l", xlab = "time", ylab = "f(t)")
  abline(h = 0, lty = 3)
```

This could then be further enhanced by using a function to create our fourier
function.

``` {r fourierfnfn}
# Define a function that gives a function as an output
  fourierfn <- function(freq, phase, amp) {
    function(t, w) {
      sum(amp*sin(freq*w*t + phase))
    }
  }

# Create your new fourier function and plot over the top of the previous plot
  mywave <- fourierfn(in.freq, in.phase, in.amp)
  pred <- sapply(t, function(t) fn(t, w))
  plot(t, pred, type = "l", xlab = "time", ylab = "f(t)")
  pred <- sapply(t, function(t) mywave(t, w))
  lines(t, pred, type = "l", xlab = "time", ylab = "f(t)", lty = 2, col = "red")
```

### Fourier Series using Sine and Cosine Waves

While the addition of two sine waves can be used in a Fourier curve, the same
goes for having both a sine and a cosine wave. But what advantages come with this?

``` {r wave3}
  xs <- seq(-2*pi, 2*pi, pi/100)
  wave.1 <- sin(3*xs)
  wave.2 <- cos(10*xs)
  plot(xs,wave.1, type = "l", ylim = c(-1, 1))
  abline(h = 0,lty = 3)
  plot(xs,wave.2, type = "l", ylim = c(-1, 1))
  abline(h = 0, lty = 3)
  wave.3 <- 0.5 * wave.1 + 0.25 * wave.2
  plot(xs, wave.3, type = "l")
  title("Fourier Series")
  abline(h = 0,lty = 3)
  repeat.xs <- seq(-2*pi, 0, pi/100)
  wave.3.repeat <- 0.5*sin(3*repeat.xs) + 0.25*cos(10*repeat.xs)
  points(repeat.xs, wave.3.repeat, type = "l",col = "red")
  abline(h = 0,v = c(-2*pi, 0), lty = 3)
```

The equation
for these functions are below:
$$h_1(f,j)=A+Bcos(\omega t)+Csin(\omega t)$$
$$h_2(f,j)=h_1(f,j)+Dcos(2\omega t)+Esin(2\omega t)$$
$$h_3(f,j)=h_2(f,j)+Dcos(4\omega t)+Esin(4\omega t),  etc.$$
I think these may be beneficial as they allow for phase shifting to occur 
without adding additional parameters. This would make it more beneficial for 
curve fitting.

### Use in Pharmacokinetics

Fourier Series have more components that just this and can get quite complex,
but the description above is sufficient for fitting the a Fourier curve to
pharmacokinetic data.

Lets observe the following dataset:

```{r simplot}
  with(subdata, plot(time, dv))
  with(absdata, lines(time, sumexp))
```

This data represents a drug with one-compartment pharmacokinetics and 
first-order absorption. Theoretically we can fit a Fourier curve to this!

Below is some code that fits a Fourier series to data

``` {r curvefit}
  y <- absdata$sumexp
  pr <- 2 # number of pairs
  n <- length(y) # sample size
  tmp <- rep(NA, 2*pr*n) #construction of a vector to be reshaped to a matrix
  col_p <- matrix(tmp, n, 2*pr) #reshape the vector tmp to a matrix with
  # n rows and 2*pr columns
  for(k in 1:pr) { # create pr columns of cosines followed by
    # pr columns of sines
    for(j in 1:n) {
      col_p[j,k] <- cos(2*pi*j*k/24)
      col_p[j, pr+k] <- sin(2*pi*j*k/24)
    }
  }
  ones <- rep(1,n)
  X <- cbind(ones,col_p)
  fit <- lm(y ~ 0 + X)
  with(absdata, plot(time, sumexp))
  lines(0:24, X%*%fit[[1]])
  sum(fit$resid^2) # print out SSE for the current model
```

For a fitted curve with only 1 pair the coefficients that result from `lm()` are
$A$ , $B$ and $C$ from the equation on Fourier Series for Sine and Cosine waves.

However this code only handles uniformly distributed data, and not for data taken
at times often seen in pharmacokinetic studies.

``` {r pkfit}
  with(subdata, plot(time, dv))
  t <- subdata$time
  y <- subdata$dv
  pr <- 3 # number of pairs
  n <- length(y) # sample size
  tmp <- rep(NA, 2*pr*n) #construction of a vector to be reshaped to a matrix
  col_p <- matrix(tmp, n, 2*pr) #reshape the vector tmp to a matrix with
  # n rows and 2*pr columns
  for(k in 1:pr) { # create pr columns of cosines followed by
    # pr columns of sines
    for(j in 1:n) {
      col_p[j,k] <- cos(2*pi*t[j]*k/24)
      col_p[j, pr+k] <- sin(2*pi*t[j]*k/24)
    }
  }
  ones <- rep(1,n)
  X <- cbind(ones,col_p)
  fit <- lm(y ~ 0 + X)
  with(subdata, plot(time, dv))
  lines(subdata$time, X%*%fit[[1]])
```

Using this fitting process we can then find the log-likelihood and therefore
the AIC for each model, finding the best number of pairs. In this case the 
saturated model is chosen (largest number of parameters possible).

``` {r loopfit}
  coef <- unname(fit$coefficients)
  yhat <- (X%*%coef)[,1]
  loglik <- dnorm(y, yhat, sigma, log = T)
  out <- 2*length(coef)-2*sum(loglik)
```