---
title: "Single tissue models in MoBi"
author: "Jim Hughes"
date: "Monday, 18 June 2018"
output:
  html_document: default
  html_notebook: default
---

```{r setup, include = FALSE}
# knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)
# Identify git directory and remove previous objects if not already done
  rm(list=ls(all = TRUE))
  graphics.off()
  #git.dir <- "E:/Hughes/Git"
  git.dir <- "C:/Users/Jim Hughes/Documents/GitRepos/"
  reponame <- "len_pbpk"

# Load libraries
  library(readxl)
  library(stringr)
  library(plyr)
  library(reshape2)
  library(ggplot2)
  library(knitr)

# Customize ggplot2 theme
  theme_bw2 <- theme_set(theme_bw(base_size = 14))
  theme_update(plot.title = element_text(hjust = 0.5))
  
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
# Tidy Observed Data
# Set up environment with observed data present
  source(paste(git.dir, reponame, "rscripts",
    "data_iv.R", sep = "/"))
  
# Grab plasma data
# Convert data to units PKSim is expecting
# Units are ng/mL aka ug/L
# convert to umol/L as when multiplied by volume in the arterial blood it gives
# umol which is PKSim's native unit
  dataiv.pla <- with(dataiv[!is.na(dataiv$PLA),], data.frame(
    TADNOM = TADNOM,
    DOSEMGKG = DOSEMGKG,
    TIME = TIME,
    DV = PLA/259.26,
    DVNORM = PLA/(DOSEMG*259.26)
  ))

  plasma.data <- ddply(dataiv.pla, .(TADNOM, DOSEMGKG), function(x) {
    with(x, data.frame(
      TIME = mean(TIME, na.rm = T),
      DV = mean(DV, na.rm = T),
      DVNORM = mean(DVNORM, na.rm = T)
    ))
  })
  
  forcingFunction <- function(x) {
    0.03*(
      exp(-2.286E-5*x + -1.33848) + 
      exp(-0.01448337*x + 3.024384) + 
      exp(-0.1014477*x + 5.992029)
    )
  }
  sumexp.data <- data.frame(
    TIME = rep(seq(0, 600, by = 1), times = 4),
    DOSEMGKG = rep(c(0.5, 1.5, 5, 10), each = 601)
  )
  sumexp.data$PRED <- with(sumexp.data, DOSEMGKG*forcingFunction(TIME))
   
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
# Tidy Simulation Data
# Simulation data is output from MoBi
# Time units are hours
# Concentration units are umol/L
# Molar mass of lenalidomide = 259.26 g/mol
# Read in simulation data
  file.dir <- "/raw_data/MoBi_paper/"
  model.name <- "BrainSumexp52.xlsx"
  braindata_raw <- as.data.frame(read_excel(paste0(git.dir, reponame, file.dir, model.name)))
  
# Remove hydrolysis metabolite columns
  braindata <- braindata_raw[!str_detect(names(braindata_raw), "Hydrolysis.Metabolite")]
  
# Rename columns
  names(braindata) <- c(
    "TIME", "Venous Blood Plasma", "Venous Blood Blood Cells", 
    "Arterial Blood Plasma", "Arterial Blood Blood Cells", "Tissue Plasma", 
    "Tissue Blood Cells", "Tissue Interstitial", "Tissue Intracellular", "Whole Tissue"
  )
  
# Read in simulation data
  model.name <- "HeartSumexp52.xlsx"
  heartdata_raw <- as.data.frame(read_excel(paste0(git.dir, reponame, file.dir, model.name)))
  
# Remove hydrolysis metabolite columns
  heartdata <- heartdata_raw[!str_detect(names(heartdata_raw), "Hydrolysis.Metabolite")]
  
# Rename columns
  names(heartdata) <- c(
    "TIME", "Venous Blood Plasma", "Venous Blood Blood Cells", 
    "Arterial Blood Plasma", "Arterial Blood Blood Cells", "Tissue Plasma", 
    "Tissue Blood Cells", "Tissue Interstitial", "Tissue Intracellular", "Whole Tissue"
  )
  
# Read in simulation data
  model.name <- "KidneySumexp52.xlsx"
  kidneydata_raw <- as.data.frame(read_excel(paste0(git.dir, reponame, file.dir, model.name)))
  
# Remove hydrolysis metabolite columns
  kidneydata <- kidneydata_raw[!str_detect(names(kidneydata_raw), "Hydrolysis.Metabolite")]
  
# Rename columns
  names(kidneydata) <- c(
    "TIME", "Venous Blood Plasma", "Venous Blood Blood Cells", 
    "Arterial Blood Plasma", "Arterial Blood Blood Cells", "Tissue Plasma", 
    "Tissue Blood Cells", "Tissue Interstitial", "Tissue Intracellular", "Whole Tissue"
  )
  
# Read in simulation data
  model.name <- "LungSumexp52.xlsx"
  lungdata_raw <- as.data.frame(read_excel(paste0(git.dir, reponame, file.dir, model.name)))
  
# Remove hydrolysis metabolite columns
  lungdata <- lungdata_raw[!str_detect(names(lungdata_raw), "Hydrolysis.Metabolite")]
  
# Rename columns
  names(lungdata) <- c(
    "TIME", "Venous Blood Plasma", "Venous Blood Blood Cells", 
    "Arterial Blood Plasma", "Arterial Blood Blood Cells", "Tissue Plasma", 
    "Tissue Blood Cells", "Tissue Interstitial", "Tissue Intracellular", "Whole Tissue"
  )
  
# Add tissue column and bind together
  braindata$TISSUE <- "Brain"
  braindata$K <- 0.00017816/0.00031271
  heartdata$TISSUE <- "Heart"
  heartdata$K <- 0.00038372/0.00031271
  kidneydata$TISSUE <- "Kidney"
  kidneydata$K <- 0.0017816/0.00031271
  lungdata$TISSUE <- "Lung"
  lungdata$K <- 0.0075004/0.00071923
  simdata1 <- rbind(braindata[-1,], heartdata[-1,], kidneydata[-1,])
  simdata1$OBS <- simdata1$`Arterial Blood Plasma`
  simdata2 <- lungdata[-1,]
  simdata2$OBS <- simdata2$`Venous Blood Plasma`
  simdata <- rbind(simdata1, simdata2)
  simdata$TIME <- simdata$TIME*60
  simdata$TISSUEf <- factor(simdata$TISSUE)
  mobi.data <- simdata
```

### Base Evaluation of Method
PKSim is not designed for use with single tissue models, therefore a method was
designed to convert PKSim models for use as single tissue models using MoBi.

The observed data for the venous blood was used as the input of the model. The
equation used for this forcing function were made up of a sum of exponentials.

The sum of exponentials were fitted to dose normalised data. The forcing 
function is multiplied by weight of the mouse (W) and the dose (D) to remove
the dose normalisation. The units of the function are in umol/L as these are 
the units of concentrations used in MoBi.

$$FF(t) = D*W*(e^{-3.81*10^{-7}t - 1.34} + e^{-0.0145t+3.02} + e^{-0.101t+5.99})$$

The plot below shows the plasma concentrations for each of the four dosage ranges (0.5, 1.5, 5 and 10 mg/kg). The mean observed plasma concentrations obtained from pre-clinical studies are represented by the blue dots, while the concentrations predicted by the sum of exponentials are represented by the red line.

```{r sumexpdata, echo = F}
# Plot the sum of exponential data against the mean observed values
  p1 <- NULL
  p1 <- ggplot()
  p1 <- p1 + geom_point(aes(x = TIME, y = DV), data = plasma.data, colour = "blue")
  p1 <- p1 + geom_line(aes(x = TIME, y = PRED), data = sumexp.data, colour = "red")
  p1 <- p1 + facet_wrap(~DOSEMGKG, scales = "free_y")
  p1 <- p1 + xlab("Time (mins)")
  p1 <- p1 + scale_y_log10("Concentration (nmol/mL)")
  p1
```

These sum of exponentials were used as the concentration for the afferent
blood vessel plasma and blood cells (arterial blood for most tissues, 
venous blood for lung).

The forcing function is used to dictate the concentrations in both plasma and
blood cells, however MoBi requires the amount of drug in each compartment when
at equilibrium. Therefore the forcing function is multiplied by the volume of 
each compartment. The drug amount given to the blood cell compartment is also 
multiplied by the partition coefficient between blood cells and plasma to 
ensure that both compartments are in equilibrium during simulation.

$$C_{pls} = V_{pls}*FF(t)$$
$$C_{bc} = V_{bc}*K_{bc}*FF(t)$$

When implemented in MoBi with no corrections concentrations are not
recreated faithfully, with this phenomenon being worse in some tissues.

The plot below shows the concentrations of the afferent blood vessel in four
different tissues. The concentrations determined by the forcing function are
represented by the red line, while concentrations that resulted from MoBi
simulations are represented in green.

```{r simdata, echo = F}
# Create a subset of the sum of exponential data for the 0.5mg/kg data
  sub.sumexp.data <- sumexp.data[sumexp.data$DOSEMGKG == 0.5,]

# Plot the sum of exponential data and the simulated mobi data
  p2 <- NULL
  p2 <- ggplot()
  p2 <- p2 + geom_line(aes(x = TIME, y = OBS), data = mobi.data, colour = "green4")
  p2 <- p2 + geom_line(aes(x = TIME, y = PRED), data = sub.sumexp.data, colour = "red")
  p2 <- p2 + facet_wrap(~TISSUEf, scales = "free_y")
  p2 <- p2 + xlab("Time (mins)")
  p2 <- p2 + scale_y_log10("Concentration (nmol/mL)")
  p2
```

### Sensitivity Analysis
To determine what components of the model were causing these issues a 
sensitivity analysis was performed. This involves perturbing parameters by a 
small amount and evaluating the impact this has on pharmacokinetic metrics such
as AUC and Cmax. If the forcing function is working perfectly, you would expect
the most important parameters to be those responsible for the forcing function.

Sensitivity analysis was run at multiple simulation resolutions to determine if
this also had an impact on pharmacokinetic metrics.

The most sensitive model parameters for changes in Cmax were Afferent Blood
Vessel Volume and Tissue Blood Flow. Volume should not have impacted Cmax in
the sensitivity analysis, as the dose was linked to afferent blood volume to
ensure the Cmax was the same independent of blood vessel volume.

The most sensitive model parameters for changes in AUC were also Afferent Blood
Vessel Volume and Tissue Blood Flow. Blood Flow should not have impacted AUC in
the sensitivity analysis, as the decrease in afferent blood concentrations were
dictated by a forcing function, not by the drug flowing out of the compartment.

As these parameters effect both Cmax and AUC it is somewhat safe to assume that
Cmax and AUC are sensitive to these parameters as they affect concentrations
in the compartment.

Below shows the sensitivity indicies (Value column) for the Heart tissue.

```{r sitable, echo = F}
# Read in the transcribed sensitivity analysis data
  file.name <- "transcribed_si_data.csv"
  si.data <- read.csv(paste0(git.dir, reponame, file.dir, file.name))
  names(si.data)[1] <- "Tissue"
  
# Show a subset of this data
  kable(si.data[si.data$Tissue == "Heart",])
```

Interestingly the two parameters impacted the metrics by the same amount and 
there were no other influential parameters. The image below shows the 
sensitivity index plotted against the simulation interval for each tissue.

```{r sivsres, echo = F}
# Plot the si data against the simulation resolution
  si.subdata <- si.data[si.data$Parameter == "Blood Volume",]
  p3 <- NULL
  p3 <- ggplot()
  p3 <- p3 + geom_point(aes(x = Resolution, y = Value), data = si.subdata)
  p3 <- p3 + xlab("Simulation Resolution (min-1)")
  p3 <- p3 + ylab("Sensitivity Index")
  p3 <- p3 + facet_wrap(~Tissue, scales = "free_y")
  p3
```

There is an obvious $y = 1/x$ relationship between these two variables. The
transformed data is plotted below with the slope values. The slopes 
correspond with the elimiationn rate constant determined from Tissue Blood Flow
and Afferent Blood Volume.

```{r sivsrestrans, echo = F}
# Determine the coefficients of si ~ 1/sim.res
  si.model <- ddply(si.subdata, .(Tissue), function(df) {
    X <- as.matrix(data.frame(
      ones = rep(1, dim(df)[1]),
      inv.res = 1/df$Resolution
    ))
    y <- df$Value
    mod <- lm(y ~ 0 + X)
    data.frame(
      Intercept = mod$coefficients[1],
      Slope = mod$coefficients[2]
    )
  })
  si.model$SlopeText <- paste("m =", signif(si.model$Slope, 4))

# Plot the transformed values and the linear regression slope
  p4 <- NULL
  p4 <- ggplot()
  p4 <- p4 + geom_point(aes(x = 1/Resolution, y = Value), data = si.subdata)
  p4 <- p4 + geom_text(aes(x = 2.5, y = 7.5, label = SlopeText), data = si.model)
  p4 <- p4 + xlab("Inverse Simulation Resolution (min)")
  p4 <- p4 + ylab("Sensitivity Index")
  p4 <- p4 + facet_wrap(~Tissue)
  p4
  
# Show the elimination rate constant values
  kable(data.frame(
    Tissue = c("Brain", "Heart", "Kidney", "Lung"),
    `Tissue Blood Flow` = c(0.0001782, 0.0003837, 0.001782, 0.0075),
    `Afferent Blood Volume` = c(rep(0.0003127, 3), 0.0007192),
    k = c(0.5697, 1.227, 5.697, 10.43)
  ))
```

This means that the sensitivity index is directly proportional to the 
elimination rate constant and the simulation resolution. Therefore it also
means that concentration is dependent on these parameters.

### Simulation Method Analysis
To further explore how the elimination rate constant affected model predictions
the forcing function was implemented in a single-tissue model, where the tissue
blood flow was set to zero.

```{r arterial, echo = F}
# Read in and tidy arterial simulation data, also make subset
  model.name <- "ArterialSumexp52.xlsx"
  artery.data_raw <- as.data.frame(read_excel(
    paste0(git.dir, reponame, file.dir, model.name)
  ))
  names(artery.data_raw) <- c("TIME", "sim.pls", "sim.bc")
  artery.data <- artery.data_raw[1:20, ]
  artery.data$TIME <- artery.data$TIME*60
  
# Determine sum of exponential values for artery simulation times
  artery.data$sumexp.pls <- 0.5*forcingFunction(artery.data$TIME)
  artery.data$sumexp.bc <- 0.477865612645714*0.5*forcingFunction(artery.data$TIME)
  
  kable(head(artery.data))
```

As can be seen the values in the simulation lag behind the true values. The
equation below describes this simplified model, where $C$ is the simulated 
concentration, $FF$ is the forcing function, $t$ is time and $T$ represents the
simulation resolution.

$$C(t) = FF(t-\frac{1}{T})$$

To correct for this the $\frac{1}{T}$ was added to $t$ in the forcing function
equation.

$$FF(t) = e^{-m(t+\frac{1}{T})+c}$$

The simulations for the simplified model also inform how the forcing function
is implemented. As $C(t)$ lags behind $FF(t)$ it becomes apparent that the
order of operations during simulation of a full single-tissue model are:

1. Set $t$ to zero and begin simulation
2. Report $C(t)$ in compartment at time $t$
3. Calculate $FF(t)$ and update $C(t)$
4. Add $\frac{1}{T}$ to $t$
5. Potential loss of drug?
6. Return to step 2

It was hypothesised that the drug was lost during step 5. Therefore the
equation below was used to represent $C(t)$ in a single-tissue model, where $\epsilon$ represents the drug lost in step 5. 

$$C(t) = FF(t-\frac{1}{T})+\epsilon$$

To better understand the drug loss, it was calculated by rearranging this
equation.

$$\epsilon = C(t) - FF(t-\frac{1}{T})$$

```{r resvpred, echo = F}
# Calculate residuals using lagged forcing function
  epsilon.data <- ddply(mobi.data, .(TISSUE), function(df) {
    df$FFlag <- 0.5*forcingFunction(c(0, head(df$TIME, -1)))
    df
  })
  epsilon.data$RES <- with(epsilon.data, FFlag - OBS)
  epsilon.data$PROPRES <- with(epsilon.data, RES/FFlag)
  
# Plot residuals vs concentration
  p5 <- NULL
  p5 <- ggplot()
  p5 <- p5 + geom_line(aes(x = FFlag, y = RES), data = epsilon.data)
  p5 <- p5 + facet_wrap(~TISSUE)
  p5 <- p5 + xlab("Sumexp Concentration (umol/L)")
  p5 <- p5 + ylab("Residual (umol/L)")
  p5
```

When plotting the MoBi simulated concentrationss against the residuals we see 
an obvious relationship. This suggests the residual is proportional.

```{r properrstatsum, echo = F}
# Show mean and 90% confidence intervals of proportional error values
  epsilon.stat <- ddply(epsilon.data, .(TISSUE), function(df) {
    data.frame(
      CI90lo = signif(CI90lo(df$PROPRES), 4),
      Mean = signif(mean(df$PROPRES), 4),
      CI90hi = signif(CI90hi(df$PROPRES), 4)
    )
  })
  epsilon.stat
```

During step 5 an amount of drug is lost from the afferent blood vessel that
is dependent on the tissue blood flow rate, the blood volume and the simulation
interval. It is assumed that this is due to passive diffusion between
compartments. The differential equations for this process are:

$$\frac{dA_{pls}}{dt}=-Q*(1-HCT)*C_{pls}$$
$$\frac{dC_{pls}}{dt}=-\frac{Q*(1-HCT)*C_{pls}}{V_{pls}}$$
$$\frac{dC_{pls}}{dt}=-\frac{Q*(1-HCT)*C_{pls}}{V_{blood}*(1-HCT)}$$
$$\frac{dC_{pls}}{dt}=-\frac{Q*C_{pls}}{V_{blood}}$$
This is just a simple first order kinetic reaction, where $C_0$ is equal 
to $FF(t)$.
$$C=C_0*e^{-kt}$$
$$C_{pls}=FF(t-\frac1T)*e^{-\frac{Q}{V}\frac{1}{T}}$$
$$\epsilon=FF(t-\frac1T)*(e^{-\frac{Q}{V}\frac{1}{T}}-1)$$
To check if this is correct we can use this equation to calculate the value for
$\epsilon$ and observe if they are on the line of identity.

```{r equationcheck, echo = F}
# Calculate the residuals and plot
  epsilon.data$RESPRED <- with(epsilon.data, FFlag*(1 - exp(-K/5.2)))
  
  p8 <- NULL
  p8 <- ggplot()
  p8 <- p8 + geom_point(aes(x = RES, y = RESPRED), data = epsilon.data, 
    shape = 1, colour = "blue")
  p8 <- p8 + geom_abline(slope = 1, colour = "green4", linetype = "dashed")
  p8 <- p8 + xlab("True Residual")
  p8 <- p8 + ylab("Predicted Residual")
  p8 <- p8 + facet_wrap(~TISSUE, scales = "free")
  p8
```

The residuals are never off by more than 1%, meaning this equation describes
the drug loss accurately. Therefore the plot below represents what is occuring.

```{r normalsim, echo = F}
# Create data.frame to represent what occurs during simulation
  in.data <- data.frame(SEQ = 0:4)
  in.data$TIME <- in.data$SEQ/5.2
  in.data$FF <- 0.5*forcingFunction(in.data$TIME)
  in.data$FFlag <- 0.5*c(0, forcingFunction(head(in.data$TIME, -1)))
  in.data$C <- in.data$FFlag*exp(-3/5.2)
  in.data$FFfix <- in.data$FF/exp(-3/5.2)
  in.data$Cfix <- c(0, tail(in.data$FF, -1))
  plot.data <- ddply(in.data, .(SEQ), function(df) {
    temp.seq <- 1:9/(10*5.2)
    dfin <- df$SEQ/5.2
    out <- data.frame(
      TIME = c(dfin, dfin+1E-6, temp.seq + dfin)
    )
    out$C <- c(df$C, in.data$FF[df$SEQ+1]*exp(-3*c(1E-6, temp.seq)))
    out$Cfix <- c(df$Cfix, in.data$FFfix[df$SEQ+1]*exp(-3*c(1E-6, temp.seq)))
    out
  })
  plot.data$FF <- 0.5*forcingFunction(plot.data$TIME)
  
# Plot data
  p9 <- NULL
  p9 <- ggplot()
  p9 <- p9 + geom_point(aes(x = TIME, y = C), data = in.data, 
    shape = 4, colour = "red", size = 2)
  p9 <- p9 + geom_line(aes(x = TIME, y = C), data = plot.data, colour = "red")
  p9 <- p9 + geom_line(aes(x = TIME, y = FF), data = plot.data, 
    linetype = "dashed", colour = "black")
  p9 <- p9 + xlab("Time (min)")
  p9 <- p9 + ylab("Concentration (umol/L)")
  p9
```

### Fixing One Problem...
Considering all of this we can adjust the input to counteract the intrinsic
problems with the simulation to give correct values from the forcing function.
First we use $t+\frac1T$ in $FF(t)$ to counteract the input lag. Then we divide
the value by $e^{-\frac{Q}{V}\frac{1}{T}}$. This makes our new input $I(t)$, 
which results in $C(t)$ and $FF(t)$ being equivalent.

$$I(t)=\frac{FF(t+\frac1T)}{e^{-\frac{Q}{V}\frac{1}{T}}}$$
$$C(t)=I(t-\frac1T)*e^{-\frac{Q}{V}\frac{1}{T}}$$
$$\therefore C(t) = FF(t)$$

However this only solves the problem of the afferent concentrations being
incorrect. A single-tissue model is used to modify model parameters so that the
tissue concentrations match the observed concentrations. Therefore it doesn't
matter if the forcing function is being applied properly if it results in
incorrect predictions in the tissue.

To check if tissue concentrations were being simulated faithfully, a PKSim
simulation was made, with the afferent blood vessel concentrations being fit
with a sum of exponential equation. This was then used to create a 
single-tissue model aiming to recreate the PKSim simulation results.

In the plot below the proportional residuals are plotted against time. The red
line shows the residuals from the simulation without correction, the blue line
shows the resulisuals from the simulation with corrections. The dashed lines
represent the mean proportional residual. The black line is the line of 
identity.
```{r tissuecheck, echo = F}
# Read in PKSim data
  model.name <- "PKSim_withArterial.xlsx"
  pksim.raw <- as.data.frame(read_excel(paste0(git.dir, reponame, file.dir, model.name)))
  
# Rename columns
  pksim <- pksim.raw
  names(pksim) <- c(
    "TIME", 
    paste0(
      rep(c("pls", "bc"), times = 2), 
      rep(c("Venous", "Arterial"), each = 2)
    ),
    paste0(
      rep(c("pls", "bc", "is", "ic", "tis"), times = 4), 
      rep(c("Brain", "Heart", "Kidney", "Lung"), each = 5)
    )
  )
  
# Convert time to minutes
  pksim$TIME <- pksim$TIME*60
  
# Read in MoBi data for normal Kidney simulation
  model.name <- "PKSimKidney2.xlsx"
  kidlo.raw <- as.data.frame(read_excel(paste0(git.dir, reponame, file.dir, model.name)))
  
# Rename columns
  kidlo <- kidlo.raw
  names(kidlo) <- c(
    "TIME", 
    paste0(
      c(rep(c("pls", "bc"), times = 3), "is", "ic", "ur", "tis"),
      c(rep(c("Venous", "Arterial"), each = 2), rep("Kidney", times = 6))
    )
  )
  
# Convert time to minutes
  kidlo$TIME <- kidlo$TIME*60
  kidlo$RES <- pksim$tisKidney - kidlo$tisKidney
  kidlo$PROPRES <- with(kidlo, RES/pksim$tisKidney)
  
# Read in MoBi data for fixed Kidney simulation
  model.name <- "PKSimKidneyFix2.xlsx"
  kidhi.raw <- as.data.frame(read_excel(paste0(git.dir, reponame, file.dir, model.name)))
  
# Rename columns
  kidhi <- kidhi.raw
  names(kidhi) <- c(
    "TIME", 
    paste0(
      c(rep(c("pls", "bc"), times = 3), "is", "ic", "ur", "tis"),
      c(rep(c("Venous", "Arterial"), each = 2), rep("Kidney", times = 6))
    )
  )
  
# Convert time to minutes
  kidhi$TIME <- kidhi$TIME*60
  kidhi$RES <- pksim$tisKidney - kidhi$tisKidney
  kidhi$PROPRES <- with(kidhi, RES/pksim$tisKidney)
  
# Read in MoBi data for average simulation
  model.name <- "PKSimKidneyAve.xlsx" #"PKSimKidneyEst.xlsx"
  kidmi.raw <- as.data.frame(read_excel(paste0(git.dir, reponame, file.dir, model.name)))
  
# Rename columns
  kidmi <- kidmi.raw
  names(kidmi) <- c(
    "TIME", 
    paste0(
      c(rep(c("pls", "bc"), times = 3), "is", "ic", "ur", "tis"),
      c(rep(c("Venous", "Arterial"), each = 2), rep("Kidney", times = 6))
    )
  )
  
# Convert time to minutes
  kidmi$TIME <- kidmi$TIME*60
  kidmi$RES <- pksim$tisKidney - kidmi$tisKidney
  kidmi$PROPRES <- with(kidmi, RES/pksim$tisKidney)
  
# Plot initial plot
  p0 <- NULL
  p0 <- ggplot()
  p0 <- p0 + geom_line(aes(x = TIME, y = PROPRES), data = kidlo, colour = "red")
  p0 <- p0 + geom_line(aes(x = TIME, y = PROPRES), data = kidhi, colour = "blue")
  p0 <- p0 + geom_hline(yintercept = 0, colour = "black", size = 1)
  p0 <- p0 + geom_hline(yintercept = mean(kidlo$PROPRES, na.rm = T), 
    linetype = "dashed", colour = "red")
  p0 <- p0 + geom_hline(yintercept = mean(kidhi$PROPRES, na.rm = T), 
    linetype = "dashed", colour = "blue")
  p0 <- p0 + xlab("Time (min)")
  p0 <- p0 + ylab("Proportional Residual")
  # p0 <- p0 + geom_line(aes(x = TIME, y = PROPRES), data = kidmi, colour = "green4")
  # p0 <- p0 + geom_hline(yintercept = mean(kidmi$PROPRES, na.rm = T),
  #   linetype = "dashed", colour = "green4")
  p0
```
The deviations in the proportional residual from the mean are due to the use of
an imperfect sum of exponentials used to replicate the PKSim afferent blood 
concentrations.

The corrected simulations are overpredicting tissue concentrations as they are
consistently more than the afferent blood concentrations (despite us only
being able to see the concentrations marked with an x, which seems correct).
```{r secondplot, echo = F}
  p9a <- p9
  p9a <- p9a + geom_point(aes(x = TIME, y = Cfix), data = in.data, 
    shape = 4, colour = "blue", size = 2)
  p9a <- p9a + geom_line(aes(x = TIME, y = Cfix), data = plot.data, colour = "blue")
  p9a
```

Therefore a middle ground needs to be used.

$$I(t)=\frac{FF(t+\frac1T)+\frac{FF(t+\frac1T)}{e^{-\frac{Q}{V}\frac{1}{T}}}}{2}$$
$$C(t)=I(t-\frac1T)*e^{-\frac{Q}{V}\frac{1}{T}}$$
$$\therefore C(t)=\frac{FF(t)*e^{-\frac{Q}{V}\frac{1}{T}}+FF(t)}{2}$$